Nengo: A Python tool for building large-scale functional brain models

### 2013-10-02 nengo.tex
Edits from Travis's feedback

The JavaNengo scripting interface,
described in \cite{TODO},
provided a simpler interface
to the NEF-specific portions of the simulator.
This greatly enhanced productivity,
and made it possible
to create many models with very little code.
Modellers quickly started using
the scripting interface
as their primary tool,
but still needed to access the GUI
for interactive model inspection,
and still needed to modify the simulator
to add new functionality
(e.g., to add new neuron models).
This is because Jython makes it easy
to inspect and create Java objects,
but does not make it easy to
write Java code to
inspect and create Jython objects.
Modellers, therefore, had to be
proficient in Java whether
they used the scripting interface or not.
As there are few neural modellers
that are proficient with both Java and Python syntax,
the scripting interface still lacks
convenient ways to record the results of simulations,
and to define complex experimental setups.

@@

PyNengo overcomes all of JavaNengo's limitations.
It can simulate most models
currently simulated by JavaNengo
using 11\% as many lines of code,
and depending on
only three external libraries instead of ten.
The OpenCL simulator can simulate
a large model TODO times
faster than JavaNengo's simulator.
This, along with a codebase
that is well tested and
easy to understand and extend,
provides a platform
that can simulate larger and more complex
models than Spaun,
and can therefore further
test the NEF as a theory of the brain.

@@

Ensembles and nodes can be connected together
in several ways.
Like other high-level objects,
a connection contains symbolic information
about how two objects are to be connected together.

@@

Networks can vary dramatically in size,
and can be composed of other networks.
The \texttt{Integrator} network, for example,
is composed of only one recurrently connected ensemble.
By encapsulating that logic in a network,
the purpose of that ensemble is made explicit.
On the other end of the spectrum,
the \texttt{BasalGanglia} network,
is composed of five groups of ensembles
connected with several specific functions
that together implement a ``winner take all'' circuit.
Encapsulating that logic
makes a complicated section of code
easy to include in many different models.

>>>>

The JavaNengo scripting interface,
described in \cite{TODO},
provided a simpler interface
to the NEF-specific portions of the simulator.
This greatly enhanced productivity,
and made it possible
to create many models with very little code.
Modellers quickly started using
the scripting interface
as their primary tool,
but still needed to access the GUI
for interactive model inspection,
and still needed to modify the simulator
to add new functionality
(e.g., to add new neuron models).
This is because Jython makes it easy
to inspect and create Java objects,
but does not make it easy to
write Java code to
inspect and create Jython objects.
Modellers, therefore, had to be
proficient in Java whether
they used the scripting interface or not.
As there are few neural modellers
that are proficient with both Java and Python syntax,
the scripting interface still lacks
convenient ways to record the results of simulations,
and to define complex experimental setups.
Additionally, Jython is always several versions
behind CPython, making it impossible
to use new language features
like dictionary comprehensions.
Jython also cannot access libraries
that contain any C code,
including all scientific libraries
that use NumPy.

@@

PyNengo overcomes all of JavaNengo's limitations.
It can simulate most models
currently simulated by JavaNengo
using 11\% as many lines of code,\footnote{
  JavaNengo is 40,000 lines of code, not including the GUI,
  and PyNengo is 5000 lines of code. TODO}
and depending on
only three external libraries instead of ten.
The OpenCL simulator can simulate
a large model TODO times
faster than JavaNengo's simulator.
This, along with a codebase
that is well tested and
easy to understand and extend,
provides a platform
that can simulate larger and more complex
models than Spaun,
and can therefore further
test the NEF as a theory of the brain.

@@

Ensembles and nodes can be connected together
in several ways.
Like other objects in PyNengo's object model,
a connection contains symbolic information;
in this case, a connection tracks
how two objects are to be connected together.

@@

Networks can vary dramatically in size,
and can be composed of other networks.
The \texttt{Integrator} network, for example,
is composed of only one recurrently connected ensemble.
By encapsulating that logic in a network,
the purpose of that ensemble is made explicit.
On the other end of the spectrum,
the \texttt{BasalGanglia} network
is composed of five groups of ensembles
connected with several specific functions
that together implement a ``winner-take-all'' circuit.
Encapsulating the code to create those ensembles
and connections in a network subclass
makes a complicated section of code
easy to include in many different models.
